### 2019-10-17 Manuscript/manuscript.tex
Update manuscript.tex

Laurentia refers to the craton that forms the Precambrian core of North America (Fig. \ref{fig:Laurentia_map}). Laurentia is comprised of multiple Archean provinces that had unique histories prior to their amalgamation in the Paleoproterozoic, as well as tectonic zones of crustal growth that post-date this assembly \citep{Hoffman1989a, Whitmeyer2007a}. Collision between the Superior province and the composite Slave+Rae+Hearne+Nain provinces that resulted in the Trans-Hudson orogeny represents a major event in the formation of Laurentia \citep{Corrigan2009a}. Terminal collision recorded in the Trans-Hudson orogen is estimated to have been ca. 1.86 to 1.82 Ga based on constraints such as U-Pb dating of monazite grains and zircon rims \citep[e.g.]{Skipton2016a, Weller2017a}. A period of accretionary and collision orogenesis is recorded in the constituent provinces and terranes of Laurentia leading up to the terminal collision of the Trans-Hudson orogeny. This overall story of rapid Paleoproterozoic amalgamation of Laurentia's constituent Archean provinces, including the terminal Trans-Hudson orogeny, was synthesized in the seminal \textit{United Plates of America} paper of \citet{Hoffman1988a} and has been refined in the time since -- particularly with additional geochronological constraints. Of most relevance here are the events that led to the suturing of more major Archean provinces: the Thelon orogen associated with the collision between the Slave province and the Rae province ca. 2.0 to 1.9 Ga \citep{Hoffman1989a}; the Snowbird orogen associated with ca. 1.89 Ga collision between the Rae and Hearne provinces and associated terranes \citep{Berman2007a}; the Nagssugtoqidian orogen due to the ca. 1.86 to 1.84 Ga collision between the Rae and Nain provinces \citep{St-Onge2009a}; and the Torngat orogen resulting from the ca. 1.87 to 1.85 Ga collision of the Meta Incognita province (grouped with the Rae province in older compilations) with the Nain province \citep{St-Onge2009a}. As for the Wyoming province, many models posit that it was conjoined with Hearne and associated provinces at the time of the Trans-Hudson orogeny \citep[e.g.][]{St-Onge2009a, Pehrsson2015a} or was proximal to Hearne and Superior while still undergoing continued translation up to ca. 1.80 Ga \citep{Whitmeyer2007a}. A contrasting view has been been proposed that the Wyoming province and Medicine Hat blocks was not conjoined with the other Laurentia provinces until ca. 1.72 Ga \citep{Kilian2016a}. This interpretation is argued to be consistent with geochronological constraints on monazite and metamorphic zircon indicating active collisional orogenesis associated with the Big Sky orogen on the northern margin of the craton as late as ca. 1.75 to 1.72 Ga \citep{Condit2015a} and ca. 1.72 tectonomagmatic activity in the Black Hills region \citep{Redden1990a}. However, the evidence for earlier orogenesis ca. 1.78 to 1.75 in the Black Hills \citep{Dahl1999a,Hrncir2017a}, as well as high-grade tectonism as early as ca. 1.81 Ga in the Big Sky orogen \citep{Condit2015a}, may support the interpretation of \citet{Hrncir2017a} that ca. 1.72 Ga activity is a minor overprint on ca. 1.75 terminal suturing between Wyoming and Superior. Regardless, in both of these interpretations, Wyoming is a later addition to Laurentia with final suturing post-dating ca. 1.82 Ga amalgamation of Archean provinces with the Trans-Hudson orogen further to the northeast. Overall, the collision of these Archean microcontinents between ca. 1.9 and 1.8 Ga lead to rapid amalgamation of the majority of the Laurentia craton.

@@

In this chapter, I focus on the compilation of paleomagnetic poles developed through the Nordic Paleomagnetism Workshops with some additions and modifications. The Nordic Paleomagnetism Workshops have taken the approach of using expert panels to assess paleomagnetic poles and assign them grades meant to convey the confidence that the community has in these results. While many factors associated with paleomagnetic poles can be assessed quantitatively through Fisher statistics and the precision of geochronological constraints, other aspects such as the degree to which available field tests constrain the magnetization to be primary require expert assessment. The categorizations used by the expert panel are `A' and `B' with the last panel meeting occurring in Fall 2017 in Leirubakki, Iceland. An `A' rating refers to poles that are judged to be of such high-quality that they provide essential constraints that should be satisfied in paleogeographic reconstructions. A `B' rating is associated with poles that are judged to likely provide a high-quality constraint, but have some deficiency such as remaining ambiguity in the demonstration of primary remanence or the quality/precision of available geochronologic constraints. Additional poles that were not given an `A' or `B' classification at the Nordic Workshops are referred to as not-rated (`NR'). These additional poles are taken from the Paleomagia database \citep{Veikkolainen2013a}. Many of these poles are quite valuable for reconstruction and should not be dismissed from being considered in paleogeography reconstructions. For example, there are rich data associated with intrusive lithologies of the Grenville Province that are the available paleomagnetic constraints for Laurentia at the Mesoproterozoic-Neoproterozoic boundary. However, the ages of the remanence associated with these poles is complicated by the reality that the magnetization was acquired during exhumation and such cooling ages are more difficult to robustly constrain than for eruptive units or shallow-level intrusions. As a result, the vast majority of Grenville Province poles are not given an `A' or `B' rating. However, while any one of these Grenville poles could be interpreted to suffer from temporal uncertainty, the overall preponderence of poles in a similar location at the time suggests that they need to be taken seriously within any paleogeographic reconstruction of Laurentia (although note an alternative view of an allochthonous origin discussed below).

Prior to the termination of the Trans-Hudson orogeny (before 1.8 Ga), I consider paleomagnetic poles with respect to the individual Archean provinces. There are poles in the compilation for the Slave, Wyoming, Rae, Superior and Nain provinces prior to Laurentia amalgamation. Overall, these data provide an opportunity to re-evaluate the paleomagnetic evidence for relative motions between Archean provinces prior to Laurentia assembly. A lingering question raised in \citet{Hoffman1988a} that still remains is to what extent the Archean provinces each had independent drift histories with significant separation or whether they have shared histories before experiencing fragmentation and remamalgamation. The strongest comparison in this regard comes from comparisons between paleomagnetic poles between the Superior and Slave provinces \citep{Buchan2009a, Mitchell2014a, Buchan2016a}. High-quality paleomagnetic poles from these two provinces provide strong support for differential motion between the Superior and Slave provinces between 2.2 and 1.8 Ga with the two provinces not being in their relative orientation to one another either and having distinct pole paths as constrained by 5 time periods of nearly coeval poles from 2.23 and 1.89 Ga \citep{Buchan2016a}. These data provide paleomagnetic support for the hypothesis that the Trans-Hudson orogeny is the result of closure of an ocean basin between the Superior province and the Hearne+Rae+Slave provinces as the result of differential plate tectonic motion. Following the termination of this orogeny (after 1.8 Ga), I consider poles from all the respective provinces and terranes to reflect the position of all of Laurentia.

For the Superior province, an additional complexity is that paleomagnetic poles from Siderian to Rhyacian Period (2.50 to 2.05 Ga) dike swarms, as well as deflection of dike trends, support an interpretation that there was substantial Paleoproterozoic rotation of the western Superior province relative to the eastern Superior province across the Kapuskasing Structural Zone \citep{Bates1991a, Evans2010a}. This interpretation is consistent with the hypothesis of \citet{Hoffman1988a} that the Kapuskasing Structural Zone represents major intracratonic uplift related to the Trans-Hudson orogeny. \cite{Evans2010a} propose an Euler rotation of (51\textdegree N, 85\textdegree W, -14\textdegree CCW) to reconstruct western Superior relative to eastern Superior and interpret that the rotation occurred in the time interval of 2.07 to 1.87 Ga.  I follow this interpretation and group the poles into Superior (West) and Superior (East). These poles are shown in Table 1 both in their initial reference frame and rotated into the other Superior reference frame. Uncertainty remains with respect to whether the ca. 1.88 Ga Molson dikes pole pre-dates or post-dates this rotation and thus for the time-being should be considered solely in the western Superior province reference frame.

Following Laurentia's amalgamation, poles from each part of Laurentia can be considered to reflect the position of the entire composite craton. It is worth considering the possibility that poles from zones of Paleoproterozoic and Mesoproterozoic accretion could be allochthonous to the craton. \cite{Halls2015a} argued that this was the case for late Mesoproterozoic and early Neoproterozoic poles from east of the Grenvillian allochthon boundary fault. However, the majority of researchers have considered these poles to post-date major differential motion and be associated with cooling during collapse of a thick orogenic plateau developed during continent-continent collision (e.g. \citealp{Brown2012a}). Poles with a B-rating are also included in the composite that come from Greenland, Svalbard and Scotland. These terranes were once part of contiguous Laurentia, but have subsequently rifted away. These poles need to be rotated into the Laurentia reference frame prior to use for tectonic reconstruction and I apply the  rotations shown in Table \ref{tab:terrane_rotations}.

Through the Proterozoic there are intervals where there are abundant paleomagnetic poles that constrain Laurentia's position and intervals when the record is quite sparse. To visualize the temoporal coverage of the poles and to summarize the motion, implied paleolatitudes for an interior point on Laurentia is shown in Figure \ref{fig:Laurentia_paleolatitude}. There are many high-quality paleomagnetic poles in the early Mesoproterozoic with 12  poles between ca. 1480 and 1380 from Laurentia (including Greenland poles). The best-constrained portion of the record is associated with 1110 to 1083 Ma magmatism within the Midcontinent Rift and southwest Laurentia that enables the development of well-calibrated apparent polar wander path \citep{Swanson-Hysell2019a}. The late Tonian Period also has a number of poles the Gunbarrel LIP (ca. 780 Ma) and Franklin LIP (ca. 720 Ma) as well as contemporaneous sedimentary rocks from western Laurentia basins. Overall the record of paleomagnetic poles has a lot of consistency with progressive paths and agreement between poles.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../Figures/Laurentia_reconstructions.pdf}
\caption{\small{\textbf{Paleogeogrpahic reconstructions of Laurentia at time intervals through the Proterozoic that are well-constrained by paleomagnetic data. These reconstructions use the simple Laurentia pole interpolation model that is shown in Figure \ref{fig:Laurentia_paleolatitude} and use this model to reconstruct the tectonic elements of \cite{Whitmeyer2007a} shown in Figure \ref{fig:Laurentia_map}. Modern coastlines are maintained in these polygons so that the rotated orientations can be interpreted by the reader in comparison to Figure \ref{fig:Laurentia_map}.}}}
\label{fig:Laurentia_reconstructions}
\end{figure}

@@

In this work, I develop a continuous paleogeogrpahic reconstruciton of Laurentia constrained the paleomagnetic poles compilation. This path is based on Laurentia data alone which means that it is poorly constrained through intervals of sparse data (900-800 Ma for example). One could use interpretations of paleogeographic connections with other cratons (e.g. Baltica in the Neoproterozoic) to fill in such portions of the path, however the result then becomes model-dependent without being constrained by data from Laurentia itself. The paleolatitude implied by this continuous model is shown in Figure \ref{fig:Laurentia_paleolatitude}. Paleogeographic snapshots for the past position of Laurentia are shown in Figure \ref{fig:Laurentia_reconstructions.pdf}. These reconstructions use the tectonic elements as defined by \citet{Whitmeyer2007a} with these elements being progressively added associated with Laurentia's growth. As a reminder to the reader, paleomagnetic poles provide constraints on the paleolatitude of a continental block as well as its orientation (which way was north relative to the block). While they provide constraints in this regard, they do not provide constraints in and of themselves to the longitudinal position of the block. Other approaches to obtain paleolongitude utilize geophysical hypotheses such as assuming that large low shear velocity provinces have been stable plume-generating zones in the lower mantle to which plumes can be reconstructed \citep{Torsvik2014a} or that significant pole motion in certain time intervals is associated with true polar wander axes that switch through time in conjunction with the supercontinent cycle \citep{Mitchell2012a}. In Figure \ref{fig:Laurentia_reconstructions.pdf}, Laurentia is centered on the longitudinal position of Duluth with the orientation and paleolatitude being constrained by the paleomagnetic pole compilation as synthesized in the simple Laurentia pole interpolation model.

@@

Many participants in the Nordic Paleomagnetism Workshop have contributed to the compilation and evaluation of the pole list utilized herein. Particular acknowledgement goes to David Evans for maintaining and distributing the compiled pole lists as well as additional efforts of Lauri Pesonen in maintining the Paleomagia database \citep{Veikkolainen2014a}. GPlates, and in particular the pyGPlates API, was utilized in this work \citep{Muller2018b}. Figures were made using Matplotlib \citep{Hunter2007a} in conjunction with cartopy \citep{Met-Office2010a} and pmagpy \citep{Tauxe2016a} within an interactive python environment \citep{Perez2007a}. This work was supported by NSF CAREER Grant 1847277 awarded to N.L.S.-H.

@@

\begin{table}[hbt]
\begin{tabular}{|l|l|l|l|p{2 in}|}
  \hline
& Euler pole & Euler pole & rotation & note and \\
Block & longitude & latitude & angle & citation \\
\hline
Greenland & -118.5 & 67.5 & -13.8 & Cenozoic separation of Greenland from Laurentia associated with opening of Baffin Bay and the Labrador Sea \citep{Roest1989} \\
\hline
Scotland & 161.9 & 78.6 & -31.0 & Reconstructing Atlantic opening following \cite{Torsvik2017a} \\
\hline
Svalbard & 125.0 & -81.0 & 68 & Rotate Svalbard to Laurentia in fit that works well with East Greenland basin according to \cite{Maloof2006a}\\
\hline
\end{tabular}
\caption{}
\end{table}

>>>>

Laurentia refers to the craton that forms the Precambrian core of North America (Fig. \ref{fig:Laurentia_map}). Laurentia is comprised of multiple Archean provinces that had unique histories prior to their amalgamation in the Paleoproterozoic, as well as tectonic zones of crustal growth that post-date this assembly \citep{Hoffman1989a, Whitmeyer2007a}. Collision between the Superior province and the composite Slave+Rae+Hearne+Nain provinces that resulted in the Trans-Hudson orogeny represents a major event in the formation of Laurentia \citep{Corrigan2009a}. Terminal collision recorded in the Trans-Hudson orogen is estimated to have been ca. 1.86 to 1.82 Ga based on constraints such as U-Pb dating of monazite grains and zircon rims \citep[e.g.]{Skipton2016a, Weller2017a}. A period of accretionary and collision orogenesis is recorded in the constituent provinces and terranes of Laurentia leading up to the terminal collision of the Trans-Hudson orogeny. This overall story of rapid Paleoproterozoic amalgamation of Laurentia's constituent Archean provinces, including the terminal Trans-Hudson orogeny, was synthesized in the seminal \textit{United Plates of America} paper of \citet{Hoffman1988a} and has been refined in the time since -- particularly with additional geochronological constraints. Of most relevance here are the events that led to the suturing of more major Archean provinces: the Thelon orogen associated with the collision between the Slave province and the Rae province ca. 2.0 to 1.9 Ga \citep{Hoffman1989a}; the Snowbird orogen associated with ca. 1.89 Ga collision between the Rae and Hearne provinces and associated terranes \citep{Berman2007a}; the Nagssugtoqidian orogen due to the ca. 1.86 to 1.84 Ga collision between the Rae and Nain provinces \citep{St-Onge2009a}; and the Torngat orogen resulting from the ca. 1.87 to 1.85 Ga collision of the Meta Incognita province (grouped with the Rae province in older compilations) with the Nain province \citep{St-Onge2009a}. As for the Wyoming province, many models posit that it was conjoined with Hearne and associated provinces at the time of the Trans-Hudson orogeny \citep[e.g.][]{St-Onge2009a, Pehrsson2015a} or was proximal to Hearne and Superior while still undergoing continued translation up to ca. 1.80 Ga \citep{Whitmeyer2007a}. A contrasting view has been been proposed that the Wyoming province and Medicine Hat blocks was not conjoined with the other Laurentia provinces until ca. 1.72 Ga \citep{Kilian2016b}}. This interpretation is argued to be consistent with geochronological constraints on monazite and metamorphic zircon indicating active collisional orogenesis associated with the Big Sky orogen on the northern margin of the craton as late as ca. 1.75 to 1.72 Ga \citep{Condit2015a} and ca. 1.72 tectonomagmatic activity in the Black Hills region \citep{Redden1990a}. However, the evidence for earlier orogenesis ca. 1.78 to 1.75 in the Black Hills \citep{Dahl1999a,Hrncir2017a}, as well as high-grade tectonism as early as ca. 1.81 Ga in the Big Sky orogen \citep{Condit2015a}, may support the interpretation of \citet{Hrncir2017a} that ca. 1.72 Ga activity is a minor overprint on ca. 1.75 terminal suturing between Wyoming and Superior. Regardless, in both of these interpretations, Wyoming is a later addition to Laurentia with final suturing post-dating ca. 1.82 Ga amalgamation of Archean provinces with the Trans-Hudson orogen further to the northeast. Overall, the collision of these Archean microcontinents between ca. 1.9 and 1.8 Ga lead to rapid amalgamation of the majority of the Laurentia craton.

@@

In this chapter, I focus on the compilation of paleomagnetic poles developed through the Nordic Paleomagnetism Workshops with some additions and modifications. The Nordic Paleomagnetism Workshops have taken the approach of using expert panels to assess paleomagnetic poles and assign them grades meant to convey the confidence that the community has in these results (Evans et al., this volume). While many factors associated with paleomagnetic poles can be assessed quantitatively through Fisher statistics and the precision of geochronological constraints, other aspects such as the degree to which available field tests constrain the magnetization to be primary require expert assessment. The categorizations used by the expert panel are `A' and `B' with the last panel meeting occurring in Fall 2017 in Leirubakki, Iceland. An `A' rating refers to poles that are judged to be of such high-quality that they provide essential constraints that should be satisfied in paleogeographic reconstructions. A `B' rating is associated with poles that are judged to likely provide a high-quality constraint, but have some deficiency such as remaining ambiguity in the demonstration of primary remanence or the quality/precision of available geochronologic constraints. Additional poles that were not given an `A' or `B' classification at the Nordic Workshops are referred to as not-rated (`NR'). These additional poles are taken from the Paleomagia database \citep{Veikkolainen2013a}. Many of these poles are quite valuable for reconstruction and should not be dismissed from being considered in paleogeography reconstructions. For example, there are rich data associated with intrusive lithologies of the Grenville Province that are the available paleomagnetic constraints for Laurentia at the Mesoproterozoic-Neoproterozoic boundary. However, the ages of the remanence associated with these poles is complicated by the reality that the magnetization was acquired during exhumation and such cooling ages are more difficult to robustly constrain than for eruptive units or shallow-level intrusions. As a result, the vast majority of Grenville Province poles are not given an `A' or `B' rating. However, while any one of these Grenville poles could be interpreted to suffer from temporal uncertainty, the overall preponderance of poles in a similar location at the time suggests that they need to be taken seriously within any paleogeographic reconstruction of Laurentia (although note an alternative view of an allochthonous origin discussed below).

Prior to the termination of the Trans-Hudson orogeny (before 1.8 Ga), I consider paleomagnetic poles with respect to the individual Archean provinces. There are poles in the compilation for the Slave, Wyoming, Rae, Superior and Nain provinces prior to Laurentia amalgamation. Overall, these data provide an opportunity to re-evaluate the paleomagnetic evidence for relative motions between Archean provinces prior to Laurentia assembly. A lingering question raised in \citet{Hoffman1988a} that still remains is to what extent the Archean provinces each had independent drift histories with significant separation or whether they have shared histories before experiencing fragmentation and reamalgamation. The strongest comparison in this regard comes from comparisons between paleomagnetic poles between the Superior and Slave provinces \citep{Buchan2009a, Mitchell2014a, Buchan2016a}. High-quality paleomagnetic poles from these two provinces provide strong support for differential motion between the Superior and Slave provinces between 2.2 and 1.8 Ga with the two provinces not being in their relative orientation to one another either and having distinct pole paths as constrained by 5 time periods of nearly coeval poles from 2.23 and 1.89 Ga \citep{Buchan2016a}. These data provide paleomagnetic support for the hypothesis that the Trans-Hudson orogeny is the result of closure of an ocean basin between the Superior province and the Hearne+Rae+Slave provinces as the result of differential plate tectonic motion. Following the termination of this orogeny (after 1.8 Ga), I consider poles from all the respective provinces and terranes to reflect the position of all of Laurentia.

For the Superior province, an additional complexity is that paleomagnetic poles from Siderian to Rhyacian Period (2.50 to 2.05 Ga) dike swarms, as well as deflection of dike trends, support an interpretation that there was substantial Paleoproterozoic rotation of the western Superior province relative to the eastern Superior province across the Kapuskasing Structural Zone \citep{Bates1991a, Evans2010a}. This interpretation is consistent with the hypothesis of \citet{Hoffman1988a} that the Kapuskasing Structural Zone represents major intracratonic uplift related to the Trans-Hudson orogeny. \cite{Evans2010a} propose an Euler rotation of (51\textdegree N, 85\textdegree W, -14\textdegree CCW) to reconstruct western Superior relative to eastern Superior and interpret that the rotation occurred in the time interval of 2.07 to 1.87 Ga.  I follow this interpretation and group the poles into Superior (West) and Superior (East).  Uncertainty remains with respect to whether the ca. 1.88 Ga Molson dikes pole pre-dates or post-dates this rotation and thus for the time-being should be considered solely in the western Superior province reference frame.

%These poles are shown in Table 1 both in their initial reference frame and rotated into the other Superior reference frame. NEED TO FIGURE OUT HOW TO SHOW THIS

Following Laurentia's amalgamation, poles from each part of Laurentia can be considered to reflect the position of the entire composite craton. It is worth considering the possibility that poles from zones of Paleoproterozoic and Mesoproterozoic accretion could be allochthonous to the craton. \cite{Halls2015b}} argued that this was the case for late Mesoproterozoic and early Neoproterozoic poles from east of the Grenvillian allochthon boundary fault. However, the majority of researchers have considered these poles to post-date major differential motion and be associated with cooling during collapse of a thick orogenic plateau developed during continent-continent collision (e.g. \citealp{Brown2012a}). Poles with a B-rating are also included in the composite that come from Greenland, Svalbard and Scotland. These terranes were once part of contiguous Laurentia, but have subsequently rifted away. These poles need to be rotated into the Laurentia reference frame prior to use for tectonic reconstruction and I apply the  rotations shown in Table \ref{tab:terrane_rotations}. The Euler pole and rotation is quite well-constrained for Greenland as it is associated with recent opening of Baffin Bay and the Labrador Sea (for which the rotation of \citealp{Roest1989a} is used). The reconstruction of Scotland is associated with the opening of the Atlantic (for which the rotation of \citealp{Roest1989a} is used) which is well-constrained but has more uncertainty associated with the Euler pole than that for Greenland. The reconstruction of Svalbard is more challenging given a multi-state tectonic history. The preferred Euler of \cite{Maloof2006a} is used here that designed, in particular, to honor the high degree of similarity between Tonian sediments in East Greenland \citep{Hoffman2012a} and those of East Svalbard \citep{Maloof2006a}.

Through the Proterozoic there are intervals where there are abundant paleomagnetic poles that constrain Laurentia's position and intervals when the record is quite sparse. To visualize the temoporal coverage of the poles and to summarize the motion, implied paleolatitudes for an interior point on Laurentia is shown in Figure \ref{fig:Laurentia_paleolatitude}. The ages of poles are also shown in comparison to the simplified summary of tectonic events in Figure \ref{fig:tectonic_history}. Both collisional and extensional tectonism can result in the formation of lithologies that can be used to develop paleomagnetic poles either as a result of basin formation, magmatism or both. In addition, intraplate magmatism resulting from plume-related large-igneous provinces can lead to paleomagnetic poles in periods that are otherwise characterized by tectonic quisence (e.g. the ca. 1267 Ma Mackenzie LIP; Fig. \ref{fig:tectonic_history}. In particular, there are many high-quality paleomagnetic poles in the early Mesoproterozoic with 12  poles between ca. 1480 and 1380 from Laurentia (including Greenland poles). The best-constrained portion of the record is associated with ca. 1110 to 1083 Ma magmatism within the Midcontinent Rift and southwest Laurentia that enables the development of well-calibrated apparent polar wander path \citep{Swanson-Hysell2019a}. The late Tonian Period also has a number of poles including the Gunbarrel LIP (ca. 780 Ma) and Franklin LIP (ca. 720 Ma) as well as contemporaneous sedimentary rocks from western Laurentia basins. Overall the record of paleomagnetic poles has a lot of internal consistency with progressive paths and agreement between poles such as for the Keweenawan Track \citep{Swanson-Hysell2019a} and the late Tonian path \citep{Eyster2019a}.

Data from other terranes add resolution to the record. In particular, data from Greenland adds 12 poles between 1385 and 1160 Ma when there are only 4 poles from mainland Laurentia. Given that the rotation between Greenland and mainland Laurentia is well-constrained (Table \ref{tab:, these poles can be rotated into Laurentia coordinates and used for reconstruction of the entire continent.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{../Figures/Laurentia_reconstructions.pdf}
\caption{\small{\textbf{Paleogeogrpahic reconstructions of Laurentia at time intervals through the Proterozoic that are well-constrained by paleomagnetic data. These reconstructions use the simple Laurentia pole interpolation model that is shown in Figure \ref{fig:Laurentia_paleolatitude} and use this model to reconstruct the tectonic elements of \cite{Whitmeyer2007a} shown in Figure \ref{fig:Laurentia_map}. Modern coastlines are maintained in these polygons so that the rotated orientations can be interpreted by the reader in comparison to Figure \ref{fig:Laurentia_map}.}}}
\label{fig:Laurentia_reconstructions}
\end{figure} 
%ADD MORE RECONSTRUCTIONS. 1 per 100 million years perhaps with indication of which ones are well-constrained by models and

@@

In this work, I develop a continuous paleogeographic reconstruciton of Laurentia constrained the paleomagnetic poles compilation. This path is based on Laurentia data alone which means that it is poorly constrained through intervals of sparse data (900-800 Ma for example). One could use interpretations of paleogeographic connections with other cratons (e.g. Baltica in the Neoproterozoic) to fill in such portions of the path, however the result then becomes model-dependent without being constrained by data from Laurentia itself. The paleolatitude implied by this continuous model is shown in Figure \ref{fig:Laurentia_paleolatitude}. Paleogeographic snapshots for the past position of Laurentia are shown in Figure \ref{fig:Laurentia_reconstructions.pdf}. These reconstructions use the tectonic elements as defined by \citet{Whitmeyer2007a} with these elements being progressively added associated with Laurentia's growth. As a reminder to the reader, paleomagnetic poles provide constraints on the paleolatitude of a continental block as well as its orientation (which way was north relative to the block). While they provide constraints in this regard, they do not provide constraints in and of themselves to the longitudinal position of the block. Other approaches to obtain paleolongitude utilize geophysical hypotheses such as assuming that large low shear velocity provinces have been stable plume-generating zones in the lower mantle to which plumes can be reconstructed \citep{Torsvik2014a} or that significant pole motion in certain time intervals is associated with true polar wander axes that switch through time in conjunction with the supercontinent cycle \citep{Mitchell2012a}. In Figure \ref{fig:Laurentia_reconstructions}, Laurentia is centered on the longitudinal position of Duluth with the orientation and paleolatitude being constrained by the paleomagnetic pole compilation as synthesized in the simple Laurentia pole interpolation model.

\section*{Conclusion}

The paleogeographic record is rich in constraints through the Precambrian -- certainly the

@@

Many participants in the Nordic Paleomagnetism Workshop have contributed to the compilation and evaluation of the pole list utilized herein. Particular acknowledgement goes to David Evans for maintaining and distributing the compiled pole lists as well as additional efforts of Lauri Pesonen in maintaining the Paleomagia database \citep{Veikkolainen2014a}. GPlates, and in particular the pyGPlates API, was utilized in this work \citep{Muller2018b}. Figures were made using Matplotlib \citep{Hunter2007a} in conjunction with cartopy \citep{Met-Office2010a} and pmagpy \citep{Tauxe2016a} within an interactive python environment \citep{Perez2007a}. This work was supported by NSF CAREER Grant 1847277 awarded to N.L.S.-H.

@@

\begin{table}[hbt]
\begin{tabular}{|l|l|l|l|p{2 in}|}
  \hline
& Euler pole & Euler pole & rotation & note and \\
Block & longitude & latitude & angle & citation \\
\hline
Greenland & -118.5 & 67.5 & -13.8 & Cenozoic separation of Greenland from Laurentia associated with opening of Baffin Bay and the Labrador Sea \citep{Roest1989a} \\
\hline
Scotland & 161.9 & 78.6 & -31.0 & Reconstructing Atlantic opening following \cite{Torsvik2017a} \\
\hline
Svalbard & 125.0 & -81.0 & 68 & Rotate Svalbard to Laurentia in fit that works well with East Greenland basin according to \cite{Maloof2006a}\\
\hline
\end{tabular}
\caption{}
\label{tab:rotations}
\end{table}
